---
title: "Prediction Assignment Writeup"
author: "Nils Jungenfelt"
date: "Thursday, June 18, 2015"
output: html_document
---

##Getting and cleaning the data
Below is the code for downloading and cleaning the data used for this assignment. Variables with more than 80% NA values are removed, as well as some irrelevant variables like time stamps.
The data comes from this research: http://groupware.les.inf.puc-rio.br/har

```{r}
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv", "pml-training.csv")
download.file("http://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv", "pml-testing.csv")

train <- read.csv("pml-training.csv", na.strings=c("NA","#DIV/0!",""))
test <- read.csv("pml-testing.csv", na.strings=c("NA","#DIV/0!",""))
nonNAs <- colSums(is.na(train)) < 0.8*nrow(train)
train <- train[, nonNAs]
test <- test[, nonNAs]
nonRelevant <- 1:7
train <- train[, -nonRelevant]
test <- test[, -nonRelevant]
```

For the analysis we will need the following packages:

```{r warning=FALSE, message=FALSE}
library(caret)
library(rattle)
library(randomForest)
library(doParallel)
```

##Building the machine learning algorithm
We begin with splitting the data into a train set (80%) and a probe set (20%). Seed is set to 133 for reproducability.

```{r}
set.seed(133)

split <- createDataPartition(train$classe, p = 0.8, list=FALSE)
train.train <- train[split,]
train.probe<- train[-split,]
```

I use a random forest model to predict the variable "classe" in the train data set, and 8 fold cross-validation to avvoid overfitting. To speed up this computationally heavy process, all available cores are used.

```{r cache=TRUE}
cl <- makeCluster(detectCores())
registerDoParallel(cl)
fitControl <- trainControl(method = "repeatedcv", number = 8, repeats = 4, allowParallel=TRUE)
model <- train(classe ~ ., data = train.train, method = "rf", trControl = fitControl)
stopCluster(cl)
```

##Results
The final model indicates an out of sample error of 0.57%:

```{r}
model$finalModel
```

To test the model I will try to predict the data from the probe set that I created:

```{r}
pred <- predict(model, newdata = train.probe)
confusionMatrix(pred,train.probe$classe)
```

So the accuracy of the model is 99.29% according to my test.

Finally, I will try to predict the outcomes in the test set provided in the assignment:

```{r}
pred <- predict(model, newdata = test)
pred
```